---
title: "MakeMyTrip Airline Rating and Review Analysis"
author: "Avinash Singh Pundhir"
date: "November 21, 2015"
output: html_document
---

#Problem Statement
Identify the best and worst airlines using a given dataset of passenger reviews.

#Executive Summary

* We have followed the following approach for identification of best and worst airlines
- Identified weights associated with various ratings using reviews text mining.
- Calculated final ratings based on weighted and overall ratings.
- Identified top 10 airlines based on % recommendation, overall rating and weighted ratings
- Identified bottom 10 airlines based on % recommendation, overall rating and weighted ratings

* Top 2 Airlines:

- Asiana Airlines - For their consistent increase in customer satisfaction in 2014
- Eva air - For their consistent performance between 2011 to 2015

* Worst 2 Airlines:

- United Airlines - For the consistent decrease in customer satisfaction
- Vueling Airlines - For the lowest performance between 2014 to 2015

#Input Dataset
Input dataset contains 41396 observations for 18 variables. Some of the variables provided are following:

Passenger reviews of airlines. Using these reviews, identify the best and worst 2 airlines among these. 
Also, what do passengers like most about the best airlines and dislike the most about the worst airlines? 

* Loading Required packages

```{r echo=TRUE, message=FALSE}
require(dplyr)
require(ggplot2)
require(ggthemes)
require(gridExtra)
require(sentiment)
require(tm)
require(RColorBrewer)
require(wordcloud)
require(SnowballC)
require(reshape2)
require(lubridate)
require(scales)
```

##Loading and cleaning data

```{r echo=TRUE}
mmtmaster <- read.csv("AirlinePassengerReview.csv", 
                      stringsAsFactors = FALSE)
nameclean <- names(mmtmaster)
nameclean <- gsub("_", "", nameclean)
names(mmtmaster) <- tolower(nameclean)
mmtmaster <- rename(mmtmaster, recommended = recommended..1..yes..0..no.)
mmtmaster$airlinename <- tolower(mmtmaster$airlinename)
```

##Text Mining of Reviews

```{r, echo=TRUE}
flightreview <- select(mmtmaster, airlinename, review, recommended)
#Cleaning Data
#Remove punctuation
flightreview$review <- tolower(flightreview$review)
flightreview$review <- gsub("[[:punct:]]", "",flightreview$review)
#Remove numbers
flightreview$review <- gsub("[[:digit:]]", "", flightreview$review)
#Remove unnecessary spaces 
flightreview$review <- gsub("[ \t]{2,}", "", flightreview$review)
flightreview$review <- gsub("^\\s+|\\s+$", "", flightreview$review)
flightreview$review <- gsub("flight", "", flightreview$review)
flightreview$review <- gsub("flights", "", flightreview$review)
positivereview <- filter(flightreview, recommended == 1)
negativereview <- filter(flightreview, recommended == 0)
```

###Analysis on Customer Reviews to identify weight factor

```{r, echo=TRUE, cache=TRUE}
myCorpus <- Corpus(VectorSource(flightreview$review[]))
myCorpus <- tm_map(myCorpus, tolower)
myCorpus <- tm_map(myCorpus, PlainTextDocument)
myCorpus <- tm_map(myCorpus, removePunctuation)
myCorpus <- tm_map(myCorpus, removeNumbers)
myCorpus <- tm_map(myCorpus, removeWords, stopwords("english"))
myCorpus <- tm_map(myCorpus, stemDocument)
wordcloud(myCorpus, scale=c(5,0.5), max.words=100, random.order=FALSE, rot.per=0.35, use.r.layout=FALSE, colors=brewer.pal(8, "Dark2"))
```

* The following are the most important factors based on text mining of user reviews:
- Seat
- Staff/Crew
- Service
- Time
- Food/meal/drink

##Rating Calculation

```{r, echo=TRUE}
#Extracting ratings data from master
flightratings <- select(mmtmaster, airlinename, date, overallrating:recommended) %>% 
                 group_by(airlinename)
flightratings$seatcomfortrating <- 2*flightratings$seatcomfortrating
flightratings$cabinstaffrating <- 2*flightratings$cabinstaffrating
flightratings$foodbeveragesrating <- 2*flightratings$foodbeveragesrating
flightratings$inflightentertainmentrating <- 2*flightratings$inflightentertainmentrating
flightratings$groundservicerating <- 2*flightratings$groundservicerating
flightratings$wificonnectivityrating <- 2*flightratings$wificonnectivityrating
flightratings$valuemoneyrating <- 2*flightratings$valuemoneyrating
#Identifying total number of entries
countdf <- as.data.frame(table(flightratings$airlinename))
names(countdf) <- c("airlinename", "count")
summary(countdf)
#Extracting airlines for which we have more than 150 reviews i.e. 3rd quantile
countdf <- filter(countdf, count > 150)
flightratings <- filter(flightratings, airlinename %in% countdf$airlinename)
presummary <- summary(flightratings)
```

* Imputing NA values 

```{r echo=TRUE}
imval <- summarise(flightratings, 
                   meanoverall = mean(overallrating, na.rm = TRUE),
                   meanseatcomf = mean(seatcomfortrating, na.rm = TRUE),
                   meancabinstaff = mean(cabinstaffrating, na.rm = TRUE),
                   meanfoodbev = mean(foodbeveragesrating, na.rm = TRUE),
                   meanentertainment = mean(inflightentertainmentrating, na.rm = TRUE),
                   meangroundser = mean(groundservicerating, na.rm = TRUE),
                   meanwifi = mean(wificonnectivityrating, na.rm = TRUE),
                   meanvaluemoney = mean(valuemoneyrating, na.rm = TRUE),
                   sumrecommended = sum(recommended, na.rm = TRUE))

for(i in 1:nrow(flightratings)){
        if (is.na(flightratings$overallrating[i])){
                flightratings$overallrating[i] <- imval$meanoverall[match(flightratings$airlinename[i], imval$airlinename)]
        }
        if (is.na(flightratings$seatcomfortrating[i])){
                flightratings$seatcomfortrating[i] <- imval$meanseatcomf[match(flightratings$airlinename[i], imval$airlinename)]
        }
        if (is.na(flightratings$cabinstaffrating[i])){
                flightratings$cabinstaffrating[i] <- imval$meancabinstaff[match(flightratings$airlinename[i], imval$airlinename)]
        }
        if (is.na(flightratings$foodbeveragesrating[i])){
                flightratings$foodbeveragesrating[i] <- imval$meanfoodbev[match(flightratings$airlinename[i], imval$airlinename)]
        }
        if (is.na(flightratings$inflightentertainmentrating[i])){
                flightratings$inflightentertainmentrating[i] <- imval$meanentertainment[match(flightratings$airlinename[i], imval$airlinename)]
        }
        if (is.na(flightratings$groundservicerating[i])){
                flightratings$groundservicerating[i] <- imval$meangroundser[match(flightratings$airlinename[i], imval$airlinename)]
        }
        if (is.na(flightratings$wificonnectivityrating[i])){
                flightratings$wificonnectivityrating[i] <- imval$meanwifi[match(flightratings$airlinename[i], imval$airlinename)]
        }
        if (is.na(flightratings$valuemoneyrating[i])){
                flightratings$valuemoneyrating[i] <- imval$meanvaluemoney[match(flightratings$airlinename[i], imval$airlinename)]
        }
}
flightratings[is.na(flightratings)] <- 0
postsummary <- summary(flightratings)
```

* Comparision of Summary before and after imputing data

```{r, echo=TRUE}
presummary
postsummary
```

* Here we can se that the data imputing strategy hasn't resulted in significant changes in mean on any rating data.

##Calculating Consolidated rating

```{r, echo=TRUE}
#Function call to get consolidated rating per observation
seatcomfortwt <- 0.3
cabinstaffwt <- 0.2
foodbevwt <- 0.2
entertainmentwt <- 0.1
groundservicewt <- 0.1
wifiwt <- 0.0
valuemoneywt <- 0.1

flightratings <- mutate(flightratings, wtrating = seatcomfortwt*seatcomfortrating +
                                                  cabinstaffwt*cabinstaffrating +
                                                  foodbevwt*foodbeveragesrating +
                                                  entertainmentwt*inflightentertainmentrating +
                                                  groundservicewt*groundservicerating +
                                                  wifiwt*wificonnectivityrating + 
                                                  valuemoneywt*valuemoneyrating)
```

* Normalization to create a dataset that can be compared across different number of total reviews

```{r, echo=TRUE}
ratingcompare <-  summarise(flightratings, sumoverall = sum(overallrating), 
                                           sumwtrating = sum(wtrating), 
                                           sumrecommended = sum(recommended))
for(i in 1:nrow(ratingcompare)){
                ratingcompare$reviewcount[i] <- countdf$count[match(ratingcompare$airlinename[i], countdf$airlinename)]
}

#Normalization of data
ratingcompare$sumoverall <- ratingcompare$sumoverall/countdf$count[match(ratingcompare$airlinename, countdf$airlinename)]
ratingcompare$sumwtrating <- ratingcompare$sumwtrating/countdf$count[match(ratingcompare$airlinename, countdf$airlinename)]


#Calculation of Weighted rating
ratingcompare <- mutate(ratingcompare, 
                        finalrating = 0.5*sumoverall + 0.5*sumwtrating)
```

* Box Plot to Summarize Rating Range

```{r, echo=TRUE}
#Converting wide format data into long format using reshape2
meltratingcompare <- melt(ratingcompare, id.vars = c("airlinename", 
                                                     "reviewcount", 
                                                     "sumrecommended"), 
                          variable.name = "typeofrating", value.name = "ratingvalue")
levels(meltratingcompare$typeofrating) <- c("Overall Rating", 
                                            "Weighted Rating",
                                            "Final Rating")

plot1 <- ggplot(meltratingcompare, aes(x = factor(0), 
                                   y = ratingvalue, 
                                   fill = typeofrating,
                                   alpha = 0.9))
plot1 <- plot1 + geom_boxplot() + 
         facet_wrap(~typeofrating) + 
         stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 3, fill = "white") +
         theme_economist() + 
         scale_colour_economist() +
         theme(legend.title = element_blank(),
               plot.title = element_text(size=20, face="bold", hjust = 0.5, vjust = 0.1),
               axis.title.x = element_text(size=15, face="bold", vjust=1),
               axis.title.y = element_text(size=15, face="bold", vjust=2)) + 
         labs(x = "", y = "Rating", title = "Exploratory Analysis of Calculated Ratings")

plot1
```

##Identification of Best and Worst airlines

```{r, echo=TRUE}
#Getting top 10 airlines
topfinalrating <- arrange(ratingcompare, 
                          desc(finalrating)) %>% 
                  filter(row_number() <= 10)

#Names formatting of top airlines
topfinalrating$airlinename <- c("Garuda", "Asiana", "Evaair", "Koreanair", "Singapore",
                                "Royalbrunei", "Bangkok", "Omanair", "Anaallnippon", "Qatar")

topfinalrating <- arrange(topfinalrating, airlinename) %>%
                  mutate(percent = round(sumrecommended/reviewcount*100))

melttopfinalrating <- melt(topfinalrating, 
                          id.vars = c("airlinename", 
                                      "reviewcount", 
                                      "sumrecommended",
                                      "percent"),
                          variable.name = "typeofrating",
                          value.name = "rating")

levels(melttopfinalrating$typeofrating) <- c("Overall Rating", 
                                            "Weighted Rating",
                                            "Final Rating")

melttopreview <- melt(topfinalrating,
                      id.vars = c("airlinename",
                                  "sumoverall",
                                  "sumwtrating",
                                  "finalrating",
                                  "percent"),
                      variable.name = "type",
                      value.name = "count")

melttopreview$percent[11:20] <- "  "

#Creating data for bottom 10 airlines

bottomfinalrating <- arrange(ratingcompare, 
                          desc(finalrating)) %>% 
                  filter(row_number() > 82)

#Names formatting of Bottom airlines
bottomfinalrating$airlinename <- c("Vueling", "USairways", "Allegiant", "Sunwing", "Royalair", 
                                   "American", "Frontier", "United", "Aircanada", "Spirit")

bottomfinalrating <- arrange(bottomfinalrating, airlinename) %>%
                  mutate(percent = round(sumrecommended/reviewcount*100))

meltbottomfinalrating <- melt(bottomfinalrating, 
                          id.vars = c("airlinename", 
                                      "reviewcount", 
                                      "sumrecommended",
                                      "percent"),
                          variable.name = "typeofrating",
                          value.name = "rating")
levels(meltbottomfinalrating$typeofrating) <- c("Overall Rating", 
                                                "Weighted Rating",
                                                "Final Rating")

meltbottomreview <- melt(bottomfinalrating,
                      id.vars = c("airlinename",
                                  "sumoverall",
                                  "sumwtrating",
                                  "finalrating",
                                  "percent"),
                      variable.name = "type",
                      value.name = "count")

meltbottomreview$percent[11:20] <- "  "
```

##Plotting Best and Worst Airlines Data

```{r, echo=TRUE}
plot2 <- ggplot(melttopfinalrating, 
            aes(x = airlinename, 
                y = rating, 
                fill = typeofrating),
            alpha = 0.5)
plot2 <- plot2 + geom_bar(stat = "identity", width = 0.6, position = "dodge") + 
         coord_equal() + coord_flip() + ylim(c(0, 10)) +
         theme_economist() + scale_colour_economist()
plot2 <- plot2 + theme(axis.text.y=element_text(size=14, 
                                   vjust=0.5,
                                   face = "bold"),
               legend.title = element_blank(),
               plot.title = element_text(size=20, face="bold", hjust = 0.5, vjust = 0.1),
               axis.title.x = element_text(size=15, face="bold", vjust=1),
               axis.title.y = element_text(size=15, face="bold", vjust=2)) +
        labs(x = "Airline Name", y = "Ratings", title = "Top Rated Airlines")
plot2


plot3 <- ggplot(data = melttopreview, aes(x = airlinename, y = count, label = percent)) 
plot3 <- plot3 + geom_point(size = 4, color = "red") + geom_line(size = 2, color = "gray") +             theme_economist() + scale_colour_economist()
plot3 <- plot3 + theme(axis.text.x=element_text(angle=90, 
                                   size=14, 
                                   vjust=0.5,
                                   face = "bold"),
           plot.title = element_text(size=20, face="bold", hjust = 0.5, vjust = 0.1),
               axis.title.x = element_text(size=15, face="bold", vjust=1),
               axis.title.y = element_text(size=15, face="bold", vjust=2)) +
        labs(x = "Airline Name", y = "Count", title = "Total Reviews vs Recommended \n(% Recommended) Top 10 Airlines") + 
        geom_text(hjust = -0.5, vjust = -2, face = "bold") 

plot3
#Bottom Plots
plot4 <- ggplot(meltbottomfinalrating, 
            aes(x = airlinename, 
                y = rating, 
                fill = typeofrating),
            alpha = 0.5)
plot4 <- plot4 + geom_bar(stat = "identity", width = 0.6, position = "dodge") + 
         coord_equal() + coord_flip() + ylim(c(0, 10)) +
         theme_economist() + scale_colour_economist()
plot4 <- plot4 + theme(axis.text.y=element_text(size=14, 
                                   vjust=0.5,
                                   face = "bold"),
               legend.title = element_blank(),
               plot.title = element_text(size=20, face="bold", hjust = 0.5, vjust = 0.1),
               axis.title.x = element_text(size=15, face="bold", vjust=1),
               axis.title.y = element_text(size=15, face="bold", vjust=2)) +
        labs(x = "Airline Name", y = "Ratings", title = "Lowest Rated Airlines")

plot4

plot5 <- ggplot(data = meltbottomreview, aes(x = airlinename, y = count, label = percent)) 
plot5 <- plot5 + geom_point(size = 4, color = "red") + geom_line(size = 2, color = "gray") +             theme_economist() + scale_colour_economist()
plot5 <- plot5 + theme(axis.text.x=element_text(angle=90, 
                                   size=14, 
                                   vjust=0.5,
                                   face = "bold"),
           plot.title = element_text(size=20, face="bold", hjust = 0.5, vjust = 0.1),
               axis.title.x = element_text(size=15, face="bold", vjust=1),
               axis.title.y = element_text(size=15, face="bold", vjust=2)) +
        labs(x = "Airline Name", y = "Count", title = "Total Reviews vs Recommended \n(% Recommended) Bottom 10 Airlines") + 
        geom_text(hjust = -0.5, vjust = -2, face = "bold") 

plot5
```

* Based on the above analysis the following are the findings:
* Top 5 Airlines
- Asiana
- Garuda
- Singapore
- Royalbrunei
- Evaair
* Bottom 5 Airlines
- Vuellng
- American
- Sumwing
- Aircanada
- United

##Identification of Top 2 Airlines

```{r, echo=TRUE}
#Extracting data for top 5 and bottom 5 airlines from master data
topfivedata <- filter(flightratings, airlinename %in% c("asiana-airlines",
                                                        "garuda-indonesia",
                                                        "singapore-airlines",
                                                        "royal-brunei-airlines",
                                                        "eva-air"))

topfivedata$airlinename <- as.factor(topfivedata$airlinename)
topfivedata$date <- dmy(topfivedata$date)
topfivedata$date <- floor_date(topfivedata$date, unit = "year")
#topfivedata$date <- quarter(topfivedata$date, with_year = TRUE)
topfivedata <- group_by(topfivedata, airlinename, date) %>%
                summarise(wtratingmean = mean(wtrating), 
                          overallratingmean = mean(overallrating)) %>%
                mutate(rateofchage = (wtratingmean/lag(wtratingmean - 1) -1 )*100)

plot6 <- ggplot(topfivedata, aes(x = date, y = rateofchage, color = airlinename))
plot6 <- plot6 + geom_line(lwd = 0.8)
plot6 <- plot6 + theme(axis.text.x=element_text(size=14, vjust=0.5, face = "bold"), 
                       plot.title = element_text(size=20, face="bold", hjust = 0.5, vjust = 0.8),
               axis.title.x = element_text(size=15, face="bold", vjust=1),
               axis.title.y = element_text(size=15, face="bold", vjust=2),
               legend.title = element_blank()) +
        labs(x = "Year", y = "% Change in Rating", title = "Change in Rating Over Time\nTop 5 Airlines")

plot6
```

* Based on the changes in weighted ratings over time the following are the top two airlines:
- Asiana Airlines - For their consistent increase in customer satisfaction in 2014
- Eva air - For their consistent performance between 2011 to 2015

##Identification of Worst 2 Airlines

```{r, echo=TRUE}
bottomfivedata <- filter(flightratings, airlinename %in% c("vueling-airlines", 
                                                           "american-airlines",
                                                           "sunwing-airlines",
                                                           "air-canada-rouge",
                                                           "united-airlines"))

bottomfivedata$airlinename <- as.factor(bottomfivedata$airlinename)
bottomfivedata$date <- dmy(bottomfivedata$date)
#bottomfivedata$date <- floor_date(bottomfivedata$date, unit = "month")
bottomfivedata$date <- quarter(bottomfivedata$date, with_year = TRUE)
bottomfivedata <- filter(bottomfivedata, date >= 2014.1)
bottomfivedata <- group_by(bottomfivedata, airlinename, date) %>%
                  summarise(wtratingmean = mean(wtrating), 
                            overallratingmean = mean(overallrating)) %>%
                  mutate(rateofchage = (wtratingmean/lag(wtratingmean - 1) -1 )*100)

plot7 <- ggplot(bottomfivedata, aes(x = date, y = rateofchage, color = airlinename))
plot7 <- plot7 + geom_line(lwd = 0.8)
plot7 <- plot7 + theme(axis.text.x=element_text(size=14, vjust=0.5, face = "bold"), 
                       plot.title = element_text(size=20, face="bold", hjust = 0.5, vjust = 0.8),
               axis.title.x = element_text(size=15, face="bold", vjust=1),
               axis.title.y = element_text(size=15, face="bold", vjust=2),
               legend.title = element_blank()) +
        labs(x = "Quarter", y = "% Change in Rating", title = "Change in Rating Over Time\nBottom 5 Airlines")

plot7
```

* Based on the changes in weighted ratings over time the following are the worst two airlines:
- United Airlines - For the consistent decrease in customer satisfaction
- Vueling Airlines - For the lowest performance between 2014 to 2015

##Finding top 2 most liked features of top two airlines

```{r, echo=TRUE, warning=FALSE}
top2data <- filter(mmtmaster, airlinename %in% c("asiana-airlines",
                                                 "eva-air")) %>%
            select(airlinename, review, recommended)
reviewtext <- paste(top2data$review, collapse = " ")
reviewtext <- gsub("flight", "", reviewtext)
reviewtext <- gsub("good", "", reviewtext)
reviewtext <- gsub("airline", "", reviewtext)
reviewtext <- gsub("asiana", "", reviewtext)
#Setting up source and corpus
reviewsource <- VectorSource(reviewtext)
topcorpus <- Corpus(reviewsource)
#Cleaning
topcorpus <- tm_map(topcorpus, tolower)
topcorpus <- tm_map(topcorpus, removePunctuation)
topcorpus <- tm_map(topcorpus, removeNumbers)
topcorpus <- tm_map(topcorpus, stripWhitespace)
topcorpus <- tm_map(topcorpus, removeWords, stopwords("english"))
#Making a Document Term Matrix
topcorpus <- tm_map(topcorpus, PlainTextDocument)
dtm <- DocumentTermMatrix(topcorpus)
dtm2 <- as.matrix(dtm)
#Finding most frequent terms
topfrequency <- colSums(dtm2)
topfrequency <- sort(topfrequency, decreasing = TRUE)
topfeatures <- head(topfrequency, 20)
wordcloud(topcorpus, scale=c(5,0.5), max.words=100, random.order=FALSE, rot.per=0.35, use.r.layout=FALSE, colors=brewer.pal(8, "Dark2"))
topfeatures
```

* The most admired features of top two airlines are following:
- Comfortable sitting arrangement
- Service provided by crew members

##Finding top 2 least liked features of worst two airlines

```{r, echo=TRUE, warning=FALSE}
bottom2data <- filter(mmtmaster, airlinename %in% c("american-airlines",
                                                 "vueling-airlines")) %>%
               select(airlinename, review, recommended)
reviewtext <- paste(bottom2data$review, collapse = " ")
reviewtext <- gsub("flight", "", reviewtext)
reviewtext <- gsub("good", "", reviewtext)
reviewtext <- gsub("airline", "", reviewtext)
#Setting up source and corpus
reviewsource <- VectorSource(reviewtext)
bottomcorpus <- Corpus(reviewsource)
#Cleaning
bottomcorpus <- tm_map(bottomcorpus, tolower)
bottomcorpus <- tm_map(bottomcorpus, removePunctuation)
bottomcorpus <- tm_map(bottomcorpus, removeNumbers)
bottomcorpus <- tm_map(bottomcorpus, stripWhitespace)
bottomcorpus <- tm_map(bottomcorpus, removeWords, stopwords("english"))
#Making a Document Term Matrix
bottomcorpus <- tm_map(bottomcorpus, PlainTextDocument)
dtm <- DocumentTermMatrix(bottomcorpus)
dtm2 <- as.matrix(dtm)
#Finding most frequent terms
bottomfrequency <- colSums(dtm2)
bottomfrequency <- sort(bottomfrequency, decreasing = TRUE)
bottomfeatures <- head(bottomfrequency, 20)
wordcloud(bottomcorpus, scale=c(5,0.5), max.words=100, random.order=FALSE, rot.per=0.35, use.r.layout=FALSE, colors=brewer.pal(8, "Dark2"))
bottomfeatures
```

* Most disliked features of bottom two airlines are following:
- Service not on time
- Seat quality

###Reference

* Informative blog on ggplot2
http://zevross.com/blog/2014/08/04/beautiful-plotting-in-r-a-ggplot2-cheatsheet-3/
* A great introduction on text mining package
https://georeferenced.wordpress.com/2013/01/15/rwordcloud/